Convergence rate of MAP estimates \\ for the exponential family

### 2021-02-23 main.tex
Change section layouts

\usepackage[round]{natbib}
\renewcommand{\bibname}{References}
\renewcommand{\bibsection}{\subsection*{\bibname}}ni

@@

\paragraph{Motivation.} 
We do not know general convergence rates on the KL for maximum likelihood estimates of the exponential family.  We want the simplest one.
We hope  to get a new result by combining tools from statistics and optimization.

\paragraph{Exponential Family.}
The exponential family member with sufficient statistic $T$ and natural parameter $\natp$ is the model
\begin{align}
    p(X|\natp) &= \exp( \natp^\top T(X) - \logpart(\natp)) \; ,
\end{align}
with log-partition function 
\begin{align}
    \logpart(\natp) = \log \int e^{\natp^\top T(x)} dx \; .
\end{align}
Recall that $\logpart$ verifies the two following identities.
\begin{align}
    \nabla\logpart(\natp) &=  \expect[p(X|\natp)]{T(X)} =: \meanp \\
    \nabla^2 \logpart(\natp) &= \Cov_\natp[T(X)] > 0
\end{align}
where $\meanp$ is called the mean parameter.
The second identity entails that $\logpart$ is strictly-convex.

\paragraph{Conjugate Prior}
The conjugate prior for $p(X|\natp)$ is
\begin{align}
    \pi(\natp) &\propto \exp( - n_0 \bregman(\natp ; \natp_0) )
\end{align}
where $n_0$ and $\natp_0$ are (hyper)parameters of the prior.
Intuitively, $n_0$ is a number of fictive points observed from a distribution with parameter $\natp_0$.
Finally, $\bregman(\natp ; \natp_0)$ is the Bregman divergence induced by $\logpart$ between $\natp$ and $\natp_0$
\begin{align}
    \bregman (\natp ; \natp_0)
    & = \logpart(\natp) - \logpart(\natp_0) 
    - \langle \nabla \logpart(\natp_0)  , \natp - \natp_0 \rangle
\end{align}
with $\nabla \logpart(\natp_0) = \expect[\natp_0]{T(X)} =: \meanp_0$ the mean parameter associated to $\natp_0$.

\paragraph{MAP.}
The negative log-likelihood of the prior is then
\begin{align*}
    -\log \pi(\natp) = n_0 (\logpart(\natp)  - \natp^\top \meanp_0 ) + \cst
\end{align*}
Thus the joint NLL of $(x_1,\dots,x_n,\natp)$ is
\begin{align}
    -\log p(X|\natp)\pi(\natp) 
    = (n_0+n) \logpart (\natp) 
    - \theta^\top \left(n_0 \meanp_0 + \sum_{i=1}^n T(x_i) \right) + \cst \; .
\end{align}
Minimizing this expression over $\natp$ yields the Maximum A Posteriori estimate
\begin{align}
    \hat \natp = \argmin_\natp -\log p(X|\natp) + n_0 \bregman(\natp ; \natp_0)
\end{align}
such that the MAP is
\begin{align}
    \nabla \logpart(\hat \natp) = \hat \meanp
    = \frac{n_0 \meanp_0 + \sum_{i=1}^n T(X_i) }{n_0+n} \; .
\end{align}
The MAP estimate is a random quantity. We wish to bound its deviation the optimum $\natp^*$ or $\meanp^*$.

\section{Straightforward Convergence Rate}
In the realizable case, the suboptimality on the population log-likelihood is exactly the KL between our current model and the true distribution
\begin{align}
    &\expect[X\sim p(.|\natp^*)]{-\log p(X|\natp) + \log p(X|\natp^*)}\\
    =& \KL( p(.|\natp^*) ; p(.|\natp)) \\
    =& \bregman (\natp ; \natp^*) \\
    =& \bregmanconj ( \meanp^* ; \meanp) \\
    =& \conj(\mu^*) + \logpart(\natp) - \langle \natp , \mu^* \rangle
    %& = \logpart(\natp) - \logpart(\natp^*) - \meanp^{*\top}(\natp - \natp^*) \; .
\end{align}
\RLP{Add details for 11 12}
where $\conj$ is the entropy, the convex conjugate of the log-partition.
The relationship between Bregman divergences and Fenchel conjugacy is well explained in \citet{wainwright2008graphical}, and \citet{agarwal2010geometric}. 
The question is: how does this quantity behave when $\natp$ is the maximum-likelihood or the MAP estimate ? Can we get bounds -- in expectation or high-probability -- on the following quantities ?
\begin{align}
	\label{eq:bregmanMLE}
	\bregmanconj \left (\E [T(X)] ;  \inv{n}  \smallsum_i T(X_i) \right ) \; , \\
	\label{eq:bregmanMAP}
	\bregmanconj \left (\E [T(X)] ; \frac{n_0 \mu_0 + \smallsum_i T(X_i)}{n_0+n} \right ) \; .
\end{align}

@@

\paragraph{$\ell^2$-norm analysis}
Let us make these statements more precise. A bound on the variance becomes a bound on the variance of the average
\begin{align}
	\Var T(X) = \E \norm{T(X) - \mu^*}^2 = \sigma^2 \\
	\implies \E \norm{\mu^* -  \inv{n}  \smallsum_i T(x_i)}^2 = \frac{\sigma^2}{n} 
\end{align}
eg the variance of the mean is $n$ times smaller than the variance of the samples.
Adding a reference mean $\mu_0$ to get the MAP yields
\begin{align}
	\E \norm{\mu^* -  \frac{n_0 \mu_0 + \smallsum_i T(x_i)}{n_0+n} }^2 
	&= \frac{n}{(n+n_0)^2} \sigma^2 +  \frac{n_0^2}{(n+n_0)^2} \norm{\mu^* -  \mu_0}^2 \\
	&= O\left(\frac{\sigma^2}{n} \right) + O\left(\frac{\norm{\mu^* -  \mu_0}^2}{n^2} \right)
\end{align}
so we have a variance term in $O(n^{-1})$ and a bias term decreasing as $O(n^{-2})$. Let's see if we can get similar estimates for arbitrary exponential families !

In the limit, the MAP estimates reach a rate $O(1/2n)$. This is shown by approximating the Bregman divergence using a second order Taylor expansion
\begin{align}
    \bregmanconj(\mu^* ; \mu) 
    &= \half (\mu - \mu^*)^\top \nabla^2\conj(\mu^*)(\mu - \mu^*)  
    + O(\norm{\mu - \mu^*}^3)\\
    &=\half  \norm{\mu^* - \mu}^2_{\nabla^2\conj(\mu^*)}
    + O(\norm{\mu - \mu^*}^3)
\end{align}
where we introduced the Mahalanobis distance induced by the matrix 
\begin{align}
    \nabla^2\conj(\mu^*) 
    = \nabla^2\logpart(\natp^*)^{-1} 
    = \Cov_{\natp^*}[T(X)]^{-1} 
    =: \mSigma^{-1}  \; .
\end{align}
To exploit this approximation, let us extend the $\ell^2$ norm results to a Mahalanobis distance induced by matrix $\mM$
\begin{align}
	\E \norm{\mu^* -  \inv{n}  \smallsum_i T(x_i)}_\mM^2 = \inv{n} \Tr(\mM \mSigma)
\end{align}
where $\mSigma$ is the covariance of $T(X)$. We retrieve $\nicefrac{\sigma^2}{n}$, the variance divided by the number of samples, when the metric is the identity $\mM=\mI$.  For the MLE we get
\begin{align}
	\E \bregmanconj \left (\E [T(X)] ;  \inv{n}  \smallsum_i T(X_i) \right ) = \frac{d}{2n} + O(n^{- \frac{3}{2}})
\end{align}
and for the MAP we  get 
\begin{align}
	\E \norm{\mu^* -  \frac{n_0 \mu_0 + \smallsum_i T(x_i)}{n_0+n} }^2_{\mSigma^{-1}}
	&= \frac{n d}{(n+n_0)^2}  +  \frac{n_0^2}{(n+n_0)^2} \norm{\mu^* -  \mu_0}^2_{\mSigma^{-1}} \\
	&= O\left(\frac{d}{n} \right) + O\left(\frac{\norm{\mu^* -  \mu_0}^2_{\mSigma^{-1}} }{n^2} \right) \; .
\end{align}
Remark how the variance does not even appear, only the dimension divided by $n$ really matters. I find this result quite spectacular : the convergence speed of MLE is not affected by the covariance of sufficient statistics, and for MAP it matters only in the $O(n^{-2})$ term.
Indeed this is because it is hidden within the Bregman divergence loss. 
Anyway that's all good for asymptotic results, but we are interested in a finite sample analysis. How do we get this ?

\subsection{Cramer-Rao Lower Bound}
In statistics, there is this famous theorem known as the Cramer-Rao lower bound. In its simplest form, it states that the covariance of an unbiased estimator of a distribution's parameter $\natp$ is lower bounded by the inverse Fisher information matrix, divided by $n$
\begin{align}
    \Cov(\hat \natp_n)  
    &\succeq \inv{n} \cI(\natp_*)^{-1} \\
    \iff
    \expect[\cD\sim \natp_*]{(\hat \natp_n - \natp_*)(\hat \natp_n - \natp_*)^T}
    &\succeq \inv{n} \expect[X\sim \natp_*]{-\nabla^2 \log p(X|\natp_*)}^{-1} \; .
\end{align}
This is a typical statistic bound, which is concerned with the $\ell^2$ accuracy of an estimator, via its covariance. In contrast, here we are concerned with the quality of the estimator in terms of KL-divergence, eg "how accurately am I modeling this distribution?"

\subsection{Posterior Expected Loss}

\section{Stochastic Bregman Proximal Point}
We see that the MAP estimate minimizes the sum of a stochastic loss $ -\log p(X|\natp)$ and a deterministic divergence to an initial point $n_0 \bregman(\natp ; \natp_0)$. This is a stochastic Bregman proximal step with step-size $\inv{n_0}$. This can also be seen at each step since
\begin{align}
    \hat \natp_{n+1} 
    &= \argmin_\natp -\log p(x_{n}|\natp) + (n_0 + n) \bregman(\natp ; \hat \natp_n) \\
    &= \argmin_\natp f(\natp; x_{n}) +  \inv{\gamma_{n}}\bregman(\natp ; \hat \natp_n) \; .
\end{align}
Hence the MAP estimate can also be seen as the result of a stochastic proximal Bregman point algorithm with step-size $\gamma_n = \inv{n_0 + n}$ at step $n$.

The analysis of deterministic Bregman proximal point relies on the three points lemma (Appendix~\ref{app:3points}). First one can prove descent, then one can prove that the total path length and the sum of suboptimalities are bounded, but this does not transpose immediately to the stochastic setting. One needs an assumption on the quality of the stochastic estimates $f(. ; x)$. 
\RLP{Write down the formulas going with this text.}
Given the proof, the most straightforward such assumption is 
\begin{align}
\expect{f(\natp_{n+1}) - f(\natp_{n+1} ; x_{n})} \leq \gamma_n \sigma^2
\end{align}
which translates to $\Cov(T(x_{n}) , \natp_{n+1}) \leq  \gamma_n \sigma^2$ in our case. This is exactly the same assumption as given by the SMD analysis. However even with this assumption, the convergence rate applies to the iterate $\sum_t \lr_t \natp_{t+1} / \sum_t \lr_t$, which gives more weight to the first iterates. It does not immediately apply to the last iterate.

@@

This is very interesting to us.
Let us have a look at their variance assumption \citep[Assumption 5.1]{hanzely2018fastest}.
Let $g_t$ be the random gradient at step $t$  (coming from data point $x_t$), and $\natp_{t+1}$ the next iterate. Then the variance bound $\sigma^2$ is an upper bound on the covariance between the gradient update  $-g_t$ and the descent direction $\natp_{t+1} - \natp_t = \nabla \logpart^*(\nabla \logpart(\natp_t) - \lr_t g_t ) - \natp_t $ that should hold for all time steps  
	&\Cov(- g_t,\natp_{t+1} - \natp_t | X_{1 \dots t}) \leq \lr_t \sigma^2 \\
	&= \expect{\langle - g_t, \natp_{t+1} \rangle}  - \langle \E[-g_t], \E[\natp_{t+1}] \rangle   \\
	 &= \expect{\langle \nabla f(\natp_t) - g_t , \natp_{t+1}  \rangle}
where $\gamma_t$ is the step-size and expectations are conditional on the past. Remark that when we plug in $\conj = \|.\|^2$, we recover the gradient variance typical of SGD. In our case, 

@@

When we plug this bound into the expected suboptimality formula~\eqref{eq:bregmanMLE} with the maximum likelihood estimate $\hat \mu =  \inv{n}\sum_i T(X_i)$, assuming all $T(x)$ belong to the marginal polytope, we get

@@

\paragraph{Open questions}
The question remains of : when is such a bound valid?  what does $\sigma^2$ depend on ? To answer these questions and more, we need to turn to examples, see  Section~\ref{sec:gaussian-variance}.

@@

@@

>>>>

\usepackage[round]{natbib}
\renewcommand{\bibname}{References}
\renewcommand{\bibsection}{\subsection*{\bibname}}

@@

\paragraph{Motivation.} 
We do not know general convergence rates on the KL for maximum likelihood estimates of the exponential family.  We want the simplest one.
We hope  to get a new result by combining tools from statistics and optimization. \TODO{Introduce Gaussian variance right from the start.}

\paragraph{Exponential Family.}
The exponential family member with sufficient statistic $T$ and natural parameter $\natp$ is the model 
\begin{align}
    p(X|\natp) &= \exp( \natp^\top T(X) - \logpart(\natp)) \; ,
\end{align}
with log-partition function 
\begin{align}
    \logpart(\natp) = \log \int e^{\natp^\top T(x)} dx \; .
\end{align}
Recall that $\logpart$ verifies the two following identities.
\begin{align}
    \nabla\logpart(\natp) &=  \expect[p(X|\natp)]{T(X)} =: \meanp \\
    \nabla^2 \logpart(\natp) &= \Cov_\natp[T(X)] > 0
\end{align}
where $\meanp$ is called the mean parameter.
If the sufficient statistic $T$ is minimal, then the log-partition function $\logpart$ is strictly convex and its gradient $\nabla \logpart$ is a bijection between natural parameters $\natp$ and mean parameters $\mu$.
The second identity entails that $\logpart$ is strictly-convex.
For a full review of exponential families and their duality, see \citet[Chapter 3]{wainwright2008graphical}. \TODO{Add standard illustration}

\paragraph{Conjugate Prior}
The conjugate prior for $p(X|\natp)$ is
\begin{align}
    \pi(\natp) &\propto \exp( - n_0 \bregman(\natp ; \natp_0) )
\end{align}
where $n_0$ and $\natp_0$ are (hyper)parameters of the prior.
Intuitively, $n_0$ is a number of fictive points observed from a distribution with parameter $\natp_0$.
Finally, $\bregman(\natp ; \natp_0)$ is the Bregman divergence induced by $\logpart$ between $\natp$ and $\natp_0$
\begin{align}
    \bregman (\natp ; \natp_0)
    & = \logpart(\natp) - \logpart(\natp_0) 
    - \langle \nabla \logpart(\natp_0)  , \natp - \natp_0 \rangle
\end{align}
with $\nabla \logpart(\natp_0) = \expect[\natp_0]{T(X)} =: \meanp_0$ the mean parameter associated to $\natp_0$. We can re-write this prior as 
\begin{align}
    \pi(\natp) \propto 
    \exp( -n_0 \logpart (\natp) 
    + \langle n_0 \mu_0, \natp \rangle ) \; ,
\end{align}
which is the formula for the exponential family with sufficient statistics $(\natp , \logpart(\natp))$ and with natural parameter $(n_0 \mu_0, -n_0)$.

\paragraph{MAP.}
The negative log-likelihood of the prior is then
\begin{align*}
    -\log \pi(\natp) = n_0 (\logpart(\natp)  - \natp^\top \meanp_0 ) + \cst
\end{align*}
Thus the joint NLL of $(x_1,\dots,x_n,\natp)$ is
\begin{align}
    -\log p(X|\natp)\pi(\natp) 
    = (n_0+n) \logpart (\natp) 
    - \theta^\top \left(n_0 \meanp_0 + \sum_{i=1}^n T(x_i) \right) + \cst \; .
\end{align}
Minimizing this expression over $\natp$ yields the Maximum A Posteriori estimate
\begin{align}
    \hat \natp = \argmin_\natp -\log p(X|\natp) + n_0 \bregman(\natp ; \natp_0)
\end{align}
such that the MAP is
\begin{align}
    \nabla \logpart(\hat \natp) = \hat \meanp
    = \frac{n_0 \meanp_0 + \sum_{i=1}^n T(X_i) }{n_0+n} \; .
\end{align}
The MAP estimate is a random quantity. We wish to bound its deviation from the optimum $\natp^*$ or $\meanp^*$.

\section{Problem}
In the realizable case, the suboptimality on the population log-likelihood is exactly the KL between our current model and the true distribution
\begin{align}
    &\expect[X\sim p(.|\natp^*)]{-\log p(X|\natp) + \log p(X|\natp^*)}\\
    =& \KL( p(.|\natp^*) ; p(.|\natp)) \\
    =& \bregman (\natp ; \natp^*) \\
    =& \bregmanconj ( \meanp^* ; \meanp) \\
    =& \conj(\mu^*) + \logpart(\natp) - \langle \natp , \mu^* \rangle
    %& = \logpart(\natp) - \logpart(\natp^*) - \meanp^{*\top}(\natp - \natp^*) \; .
\end{align}
\TODO{Add details for 11 12}
where $\conj$ is the entropy, the convex conjugate of the log-partition.
The question is: how does this quantity behave when $\natp$ is the maximum-likelihood or the MAP estimate ? Can we get bounds -- in expectation or high-probability -- on the following quantities ?
\begin{align}
	\label{eq:bregmanMLE}
	\bregmanconj \left (\E [T(X)] ;  \inv{n}  \smallsum_i T(X_i) \right ) \; , \\
	\label{eq:bregmanMAP}
	\bregmanconj \left (\E [T(X)] ; \frac{n_0 \mu_0 + \smallsum_i T(X_i)}{n_0+n} \right ) \; ,
\end{align}
where the randomness is coming from the dataset $X_1, \dots, X_n$.

\section{Comparison with  Euclidean Distance}
Often we can relate this Bregman suboptimality with the $\ell^2$ distance in mean parameter space.

@@

\paragraph{$\ell^2$-norm Analysis}
Let us make these statements more precise. A bound on the variance becomes a bound on the variance of the average
\begin{align}
	\Var T(X) = \E \norm{T(X) - \mu^*}^2 = \sigma^2 \\
	\implies \E \norm{\mu^* -  \inv{n}  \smallsum_i T(x_i)}^2 = \frac{\sigma^2}{n} 
\end{align}
eg the variance of the mean is $n$ times smaller than the variance of the samples.
Adding a reference mean $\mu_0$ to get the MAP yields
\begin{align}
	\E \norm{\mu^* -  \frac{n_0 \mu_0 + \smallsum_i T(x_i)}{n_0+n} }^2 
	&= \frac{n}{(n+n_0)^2} \sigma^2 +  \frac{n_0^2}{(n+n_0)^2} \norm{\mu^* -  \mu_0}^2 \\
	&= O\left(\frac{\sigma^2}{n} \right) + O\left(\frac{\norm{\mu^* -  \mu_0}^2}{n^2} \right)
\end{align}
so we have a variance term in $O(n^{-1})$ and a bias term decreasing as $O(n^{-2})$. Let's see if we can get similar estimates for arbitrary exponential families !

\section{Asymptotic Behaviour}
In the limit, the MAP estimates reach a rate $O(1/2n)$. This is shown by approximating the Bregman divergence using a second order Taylor expansion
\begin{align}
    \bregmanconj(\mu^* ; \mu) 
    &= \half (\mu - \mu^*)^\top \nabla^2\conj(\mu^*)(\mu - \mu^*)  
    + O(\norm{\mu - \mu^*}^3)\\
    &=\half  \norm{\mu^* - \mu}^2_{\nabla^2\conj(\mu^*)}
    + O(\norm{\mu - \mu^*}^3)
\end{align}
where we introduced the Mahalanobis distance induced by the matrix 
\begin{align}
    \nabla^2\conj(\mu^*) 
    = \nabla^2\logpart(\natp^*)^{-1} 
    = \Cov_{\natp^*}[T(X)]^{-1} 
    =: \mSigma^{-1}  \; .
\end{align}
To exploit this approximation, let us extend the $\ell^2$ norm results to a Mahalanobis distance induced by matrix $\mM$
\begin{align}
	\E \norm{\mu^* -  \inv{n}  \smallsum_i T(x_i)}_\mM^2 = \inv{n} \Tr(\mM \mSigma)
\end{align}
where $\mSigma$ is the covariance of $T(X)$. We retrieve $\nicefrac{\sigma^2}{n}$, the variance divided by the number of samples, when the metric is the identity $\mM=\mI$.  For the MLE we get
\begin{align}
	\E \bregmanconj \left (\E [T(X)] ;  \inv{n}  \smallsum_i T(X_i) \right ) = \frac{d}{2n} + O(n^{- \frac{3}{2}})
\end{align}
and for the MAP we  get 
\begin{align}
	\E \norm{\mu^* -  \frac{n_0 \mu_0 + \smallsum_i T(x_i)}{n_0+n} }^2_{\mSigma^{-1}}
	&= \frac{n d}{(n+n_0)^2}  +  \frac{n_0^2}{(n+n_0)^2} \norm{\mu^* -  \mu_0}^2_{\mSigma^{-1}} \\
	&= O\left(\frac{d}{n} \right) + O\left(\frac{\norm{\mu^* -  \mu_0}^2_{\mSigma^{-1}} }{n^2} \right) \; .
\end{align}
Remark how the variance does not even appear, only the dimension divided by $n$ really matters. I find this result quite spectacular : the convergence speed of MLE is not affected by the covariance of sufficient statistics, and for MAP it matters only in the $O(n^{-2})$ term.
Indeed this is because it is hidden within the Bregman divergence loss. 
Anyway that's all good for asymptotic results, but we are interested in a finite sample analysis. How do we get this ?

@@

This is very interesting to us.
Let us have a look at their variance assumption \citep[Assumption 5.1]{hanzely2018fastest}.
Let $g_t$ be the random gradient at step $t$  (coming from data point $x_t$), and $\natp_{t+1}$ the next iterate. Then the variance bound $\sigma^2$ is an upper bound on the covariance between the gradient update  $-g_t$ and the descent direction $\natp_{t+1} - \natp_t = \nabla \logpart^*(\nabla \logpart(\natp_t) - \lr_t g_t ) - \natp_t $ that should hold for all time steps  
\begin{align}
	&\Cov(- g_t,\natp_{t+1} - \natp_t | X_{1 \dots t}) \leq \lr_t \sigma^2 \\
	&= \expect{\langle - g_t, \natp_{t+1} \rangle}  - \langle \E[-g_t], \E[\natp_{t+1}] \rangle   \\
	 &= \expect{\langle \nabla f(\natp_t) - g_t , \natp_{t+1}  \rangle}
\end{align}
where $\gamma_t$ is the step-size and expectations are conditional on the past. Remark that when we plug in $\conj = \|.\|^2$, we recover the gradient variance typical of SGD.

\subsection{Transferring SMD's covariance assumption to MAP estimates}
In our setting, 

@@

When we plug this bound into the expected suboptimality formula~\eqref{eq:bregmanMLE} with the maximum likelihood estimate $\hat \mu =  \inv{n}\sum_i T(X_i)$, assuming all $T(X)$ belong to the marginal polytope, we get

@@

\paragraph{Open questions}
The question remains of : when is such a bound valid?  what does $\sigma^2$ depend on ? To answer these questions and more, we need to turn to examples, see  Section~\ref{sec:gaussian_variance}.

@@

\section{Cramer-Rao Lower Bound}
In statistics, there is this famous theorem known as the Cramer-Rao lower bound. In its simplest form, it states that the covariance of an unbiased estimator of a distribution's parameter $\natp$ is lower bounded by the inverse Fisher information matrix, divided by $n$
\begin{align}
    \Cov(\hat \natp_n)  
    &\succeq \inv{n} \cI(\natp_*)^{-1} \\
    \iff
    \expect[\cD\sim \natp_*]{(\hat \natp_n - \natp_*)(\hat \natp_n - \natp_*)^T}
    &\succeq \inv{n} \expect[X\sim \natp_*]{-\nabla^2 \log p(X|\natp_*)}^{-1} \; .
\end{align}
This is a typical statistic bound, which is concerned with the $\ell^2$ accuracy of an estimator, via its covariance. In contrast, here we are concerned with the quality of the estimator in terms of KL-divergence, eg "how accurately am I modeling this distribution?"

\section{Posterior Expected Loss}

@@

\section{Stochastic Bregman Proximal Point}
Remark that the MAP estimate minimizes the sum of a stochastic loss $ -\log p(X|\natp)$ and a deterministic divergence to an initial point $n_0 \bregman(\natp ; \natp_0)$. This is a stochastic Bregman proximal step with step-size $\inv{n_0}$. This can also be seen at each step since
\begin{align}
    \hat \natp_{n+1} 
    &= \argmin_\natp -\log p(x_{n}|\natp) + (n_0 + n) \bregman(\natp ; \hat \natp_n) \\
    &= \argmin_\natp f(\natp; x_{n}) +  \inv{\gamma_{n}}\bregman(\natp ; \hat \natp_n) \; .
\end{align}
Hence the MAP estimate can also be seen as the result of a stochastic proximal Bregman point algorithm with step-size $\gamma_n = \inv{n_0 + n}$ at step $n$.

The analysis of deterministic Bregman proximal point relies on the three points lemma (Appendix~\ref{app:3points}). First one can prove descent, then one can prove that the total path length and the sum of suboptimalities are bounded, but this does not transpose immediately to the stochastic setting. One needs an assumption on the quality of the stochastic estimates $f(. ; x)$. 
Given the proof, the most straightforward such assumption (see paper notes) is 
\begin{align}
\expect{f(\natp_{n+1}) - f(\natp_{n+1} ; x_{n})} \leq \gamma_n \sigma^2
\end{align}
which translates to $\Cov(T(x_{n}) , \natp_{n+1}) \leq  \gamma_n \sigma^2$ in our case. This is exactly the same assumption as given by the SMD analysis. However even with this assumption, the convergence rate applies to the iterate $\sum_t \lr_t \natp_{t+1} / \sum_t \lr_t$, which gives more weight to the first iterates. It does not immediately apply to the last iterate.
